\LARGE \textbf{Constrained Geometric Attitude Control on \( \SO \)}

### 2017-01-05 manuscript.tex
Fix typo in domain of quadratic function

\begin{prop}[Attitude Error Function] \label{prop:config_error}
%We assume a desired attitude command \( ( R_d, \Omega_d = 0 ) \), the current attitude and angular velocity \( ( R, \Omega) \), and a state constraint~\cref{eqn:constraint} are given.
Define an attitude error function \( \Psi : \SO \to \R \), an attitude error vector \( e_R \in \R^3 \), and an angular velocity error vector \( e_\Omega \in \R^3 \) as follows:
\begin{gather}
	\Psi(R) = A(R) B(R) , \label{eqn:psi} \\
	e_R = e_{R_A} B(R) + A(R) e_{R_B} , \label{eqn:eR} \\
	e_\Omega = \Omega , \label{eqn:eW}
\end{gather}
with
\begin{gather}
	A(R) = \frac{1}{2} \tr{G \left( I - R_d^T R\right)} , \label{eqn:A} \\
	B(R) = 1 - \frac{1}{\alpha} \ln \left( \frac{\cos \theta -  r^T R^T v}{1 + \cos \theta}\right) . \label{eqn:B} \\
	e_{R_A} = \frac{1}{2} \parenth{G R_d^T R - R^T R_d G} ^ \vee , \label{eqn:eRA} \\
	e_{R_B} = \frac{ \left( R^T v\right)^\wedge r}{\alpha \left(r^T R^T v - \cos \theta \right)} . \label{eqn:eRB} 
\end{gather}	
where \( \alpha \in \R \) is defined as a positive constant and the matrix \( G \in \R^{3 \times 3} \) is defined as a diagonal matrix matrix for distinct, positive constants \( g_1, g_2, g_3 \in \R \).
Then, the following properties hold
\begin{enumerate}[(i)]
	\item \label{item:prop_psi_psd} \(\Psi\) is positive definite about \( R = R_d\) on $\SO$.
	\item \label{item:prop_era}The variation of \( A(R) \) with respect to a variation of \( \delta R = R \hat{\eta} \) for \( \eta \in \R^3 \) is given by
	\begin{align}
		\dirDiff{A}{R} &= \eta \cdot e_{R_A} .
	\end{align}
	\item \label{item:prop_erb} The variation of \( B(R) \) with respect to a variation of \( \delta R = R \hat{\eta} \) for \( \eta \in \R^3 \) is given by
	\begin{align}
		\dirDiff{B}{R} &= \eta \cdot e_{R_{B}} ,
	\end{align}
  where the notation \( \D_R A \) and \(\D_R B\) represents the derivative of \( A \) or \( B \) with respect to \( R\).
    \item \label{item:prop_psi_quadratic} \( \Psi \) is locally quadratic on the domain 
    \begin{align}\label{eq:psi_quadratic_domain}
        D_1 = \braces{R \in \SO \vert \Psi = 0, e_{R} = 0 \iff R = R_d } 
    \end{align}
%	\item \label{item:prop_crit}The error function \( \Psi \) is minimized when \(R = R_d\).
	\item \label{item:prop_era_upbound}An upper bound of \( \norm{e_{R_A}} \) is given as:
	\begin{align}
		\norm{e_{R_A}}^2 \leq \frac{A(R)}{b_1} , \label{eqn:psi_lower_bound}
	\end{align}
	where the constant \( b_1 \) is given by \( b_1 = \frac{h_1}{h_2 + h_3} \) for 
	\begin{align*}
		h_1 &= \min\braces{g_1 + g_2, g_2 + g_3 , g_3 + g_1} ,\\
		h_2 &= \min\braces{\parenth{g_1 -g_2}^2,\parenth{g_2 -g_3}^2 , \parenth{g_3 -g_1}^2} ,\\
		h_3 &= \min\braces{\parenth{g_1 + g_2}^2, \parenth{g_2 + g_3}^2 , \parenth{g_3 + g_1}^2}.		
	\end{align*}
\end{enumerate}
\end{prop}
\begin{proof}
See~\Cref{proof:config_error}
\end{proof}

@@

\subsection{Proof of~\Cref{prop:adaptive_control}}\label{proof:adaptive_control}
Consider the Lyapunov function \( \mathcal{V} \) given by
\begin{align*}
	\mathcal{V} = \frac{1}{2} e_\Omega \cdot J e_\Omega + k_R \Psi + c J e_\Omega \cdot e_R + \frac{1}{2 k_\Delta} e_\Delta \cdot e_\Delta , %\label{eqn:v_adapt}
\end{align*}
over the domain \( D_3 = D_1 \cup D_2\), which ensures that the properties of~\Cref{prop:config_error,prop:error_dyn} are satisfied.

From~\Cref{prop:config_error}, the configuration error function is locally quadratic and it is bounded in \( D_3 \) by
\begin{align*}
    n_1 \norm{e_R} < \Psi <  n_2 \norm{e_{R}}^2 ,
\end{align*}
for constants \( 0 < n_1 \leq n_2 \)~\cite[Definition 6.28]{bullo2004}.
Using this the Lyapunov function \( \mathcal{V} \) is bounded by
\begin{align*} %\label{eqn:v_upper_bound}
	z^T W_1 z \leq \mathcal{V} \leq z^T W_2 z ,
\end{align*}
where \( e_\Delta = \Delta - \bar{\Delta} \), \( z = [\|e_R\|,\|e_\Omega\|,\|e_\Delta\|]^T\in\R^3 \) and the matrices \(W_1,W_2 \in \R^{3 \times 3}\) are given by
\begin{align*}
	W_1 & = \begin{bmatrix}
		k_R n_1 & \frac{1}{2} c \lambda_m & 0 \\
		\frac{1}{2} c \lambda_m & \frac{1}{2} \lambda_m & 0 \\
		0 & 0 & \frac{1}{2 k_\Delta}
	\end{bmatrix},\\
	W_2 & = \begin{bmatrix}
		k_R n_2 & \frac{1}{2} c \lambda_M & 0 \\
		\frac{1}{2} c \lambda_M & \frac{1}{2} \lambda_M & 0 \\
		0 & 0 & \frac{1}{2 k_\Delta}
	\end{bmatrix} .
\end{align*}
The time derivative of \( \mathcal{V}\) with the control input defined by~\cref{eqn:adaptive_control} is given as
\begin{align}
	\dot{\mathcal{V}} =& - k_\Omega e_\Omega^T e_\Omega + \parenth{e_\Omega + c e_R}^T W e_\Delta - k_R c e_R^T e_R \nonumber\\
	&- k_\Omega c e_R^T e_\Omega + c J e_\Omega^T \dot{e}_R - \frac{1}{k_\Delta} e_\Delta^T \dot{\bar{\Delta}} , \label{eqn:vdot}
\end{align}
where we used \( \dot{e}_\Delta = - \dot{\bar{\Delta}} \).
The terms linearly dependent on \( e_\Delta\) are combined with~\cref{eqn:delta_dot} to yield
\begin{align*}
	 e_\Delta^T \parenth{W^T \parenth{e_\Omega + c e_R} - \frac{1}{k_\Delta} \dot{\bar{\Delta}}} = 0 . 
\end{align*}
Using~\Cref{prop:eR_dot_bound} an upper bound on \( \dot{\mathcal{V}} \) is written as
\begin{gather*}
	\dot{\mathcal{V}} \leq -\zeta^T M \zeta ,
\end{gather*}
where $\zeta=[\|e_R\|,\|e_\Omega\|]\in\R^2$, and the matrix \( M \in \R^{2 \times 2} \) is 
\begin{gather*}
	M = \begin{bmatrix}
		k_R c & \frac{k_\Omega c}{2} \\
		\frac{k_\Omega c}{2} & k_\Omega - c \lambda_M H
	\end{bmatrix} .
\end{gather*}
If \( c \) is chosen such that~\cref{eqn:c_bound} is satisfied then the matrices \( W_1, W_2 \) and \( M \) are positive definite.
This implies that $\mathcal{V}$ is positive definite and decrescent, and $\dot{\mathcal{V}}$ is negative semidefinite in the domain $D_3$. As such, the zero equilibrium is stable in the sense of Lyapunov, and all of the error variables are bounded. Furthermore, $\lim_{t\to\infty} \zeta=0$ according to the LaShalle-Yoshizawa theorem~\cite{khalil1996}.

>>>>

\begin{prop}[Attitude Error Function] \label{prop:config_error}
%We assume a desired attitude command \( ( R_d, \Omega_d = 0 ) \), the current attitude and angular velocity \( ( R, \Omega) \), and a state constraint~\cref{eqn:constraint} are given.
Define an attitude error function \( \Psi : \SO \to \R \), an attitude error vector \( e_R \in \R^3 \), and an angular velocity error vector \( e_\Omega \in \R^3 \) as follows:
\begin{gather}
	\Psi(R) = A(R) B(R) , \label{eqn:psi} \\
	e_R = e_{R_A} B(R) + A(R) e_{R_B} , \label{eqn:eR} \\
	e_\Omega = \Omega , \label{eqn:eW}
\end{gather}
with
\begin{gather}
	A(R) = \frac{1}{2} \tr{G \left( I - R_d^T R\right)} , \label{eqn:A} \\
	B(R) = 1 - \frac{1}{\alpha} \ln \left( \frac{\cos \theta -  r^T R^T v}{1 + \cos \theta}\right) . \label{eqn:B} \\
	e_{R_A} = \frac{1}{2} \parenth{G R_d^T R - R^T R_d G} ^ \vee , \label{eqn:eRA} \\
	e_{R_B} = \frac{ \left( R^T v\right)^\wedge r}{\alpha \left(r^T R^T v - \cos \theta \right)} . \label{eqn:eRB} 
\end{gather}	
where \( \alpha \in \R \) is defined as a positive constant and the matrix \( G \in \R^{3 \times 3} \) is defined as a diagonal matrix matrix for distinct, positive constants \( g_1, g_2, g_3 \in \R \).
Then, the following properties hold
\begin{enumerate}[(i)]
	\item \label{item:prop_psi_psd} \(\Psi\) is positive definite about \( R = R_d\) on $\SO$.
	\item \label{item:prop_era}The variation of \( A(R) \) with respect to a variation of \( \delta R = R \hat{\eta} \) for \( \eta \in \R^3 \) is given by
	\begin{align}
		\dirDiff{A}{R} &= \eta \cdot e_{R_A} .
	\end{align}
	\item \label{item:prop_erb} The variation of \( B(R) \) with respect to a variation of \( \delta R = R \hat{\eta} \) for \( \eta \in \R^3 \) is given by
	\begin{align}
		\dirDiff{B}{R} &= \eta \cdot e_{R_{B}} ,
	\end{align}
    where the notation \( \D_R A \) and \(\D_R B\) represents the derivative of \( A \) or \( B \) with respect to \( R\).
    \item \label{item:prop_psi_quadratic} \( \Psi \) is a locally quadratic function on the domain 
    \begin{align}\label{eq:psi_quadratic_domain}
        D_1 = \braces{R \in \SO \vert \Psi = 0, e_{R} = 0 \iff R = R_d } ,
    \end{align}
    and it is bounded by constants \( 0 < n_1 \leq n_2 \) as
    \begin{align}\label{eq:psi_bound}
        n_1 \norm{e_R}^2 \leq \Psi(R) \leq n_2 \norm{e_R}^2 .
    \end{align}
%	\item \label{item:prop_crit}The error function \( \Psi \) is minimized when \(R = R_d\).
	\item \label{item:prop_era_upbound}An upper bound of \( \norm{e_{R_A}} \) is given as:
	\begin{align}
		\norm{e_{R_A}}^2 \leq \frac{A(R)}{b_1} , \label{eqn:psi_lower_bound}
	\end{align}
	where the constant \( b_1 \) is given by \( b_1 = \frac{h_1}{h_2 + h_3} \) for 
	\begin{align*}
		h_1 &= \min\braces{g_1 + g_2, g_2 + g_3 , g_3 + g_1} ,\\
		h_2 &= \min\braces{\parenth{g_1 -g_2}^2,\parenth{g_2 -g_3}^2 , \parenth{g_3 -g_1}^2} ,\\
		h_3 &= \min\braces{\parenth{g_1 + g_2}^2, \parenth{g_2 + g_3}^2 , \parenth{g_3 + g_1}^2}.		
	\end{align*}
\end{enumerate}
\end{prop}
\begin{proof}
See~\Cref{proof:config_error}
\end{proof}

@@

\subsection{Proof of~\Cref{prop:adaptive_control}}\label{proof:adaptive_control}
Consider the Lyapunov function \( \mathcal{V} \) given by
\begin{align*}
	\mathcal{V} = \frac{1}{2} e_\Omega \cdot J e_\Omega + k_R \Psi + c J e_\Omega \cdot e_R + \frac{1}{2 k_\Delta} e_\Delta \cdot e_\Delta , %\label{eqn:v_adapt}
\end{align*}
over the domain \( D_3 = D_1 \cap D_2\), which ensures that the properties of~\Cref{prop:config_error,prop:error_dyn} are satisfied.

From~\Cref{prop:config_error}, the configuration error function is locally quadratic and it is bounded in \( D_3 \) by~\cref{eq:psi_bound}~\cite[Definition 6.28]{bullo2004}.
Using this the Lyapunov function \( \mathcal{V} \) is bounded by
\begin{align*} %\label{eqn:v_upper_bound}
	z^T W_1 z \leq \mathcal{V} \leq z^T W_2 z ,
\end{align*}
where \( e_\Delta = \Delta - \bar{\Delta} \), \( z = [\|e_R\|,\|e_\Omega\|,\|e_\Delta\|]^T\in\R^3 \) and the matrices \(W_1,W_2 \in \R^{3 \times 3}\) are given by
\begin{align*}
	W_1 & = \begin{bmatrix}
		k_R n_1 & \frac{1}{2} c \lambda_m & 0 \\
		\frac{1}{2} c \lambda_m & \frac{1}{2} \lambda_m & 0 \\
		0 & 0 & \frac{1}{2 k_\Delta}
	\end{bmatrix},\\
	W_2 & = \begin{bmatrix}
		k_R n_2 & \frac{1}{2} c \lambda_M & 0 \\
		\frac{1}{2} c \lambda_M & \frac{1}{2} \lambda_M & 0 \\
		0 & 0 & \frac{1}{2 k_\Delta}
	\end{bmatrix} .
\end{align*}
The time derivative of \( \mathcal{V}\) with the control input defined by~\cref{eqn:adaptive_control} is given as
\begin{align}
	\dot{\mathcal{V}} =& - k_\Omega e_\Omega^T e_\Omega + \parenth{e_\Omega + c e_R}^T W e_\Delta - k_R c e_R^T e_R \nonumber\\
	&- k_\Omega c e_R^T e_\Omega + c J e_\Omega^T \dot{e}_R - \frac{1}{k_\Delta} e_\Delta^T \dot{\bar{\Delta}} , \label{eqn:vdot}
\end{align}
where we used \( \dot{e}_\Delta = - \dot{\bar{\Delta}} \).
The terms linearly dependent on \( e_\Delta\) are combined with~\cref{eqn:delta_dot} to yield
\begin{align*}
	 e_\Delta^T \parenth{W^T \parenth{e_\Omega + c e_R} - \frac{1}{k_\Delta} \dot{\bar{\Delta}}} = 0 . 
\end{align*}
Using~\Cref{prop:eR_dot_bound} an upper bound on \( \dot{\mathcal{V}} \) is written as
\begin{gather*}
	\dot{\mathcal{V}} \leq -\zeta^T M \zeta ,
\end{gather*}
where $\zeta=[\|e_R\|,\|e_\Omega\|]\in\R^2$, and the matrix \( M \in \R^{2 \times 2} \) is 
\begin{gather*}
	M = \begin{bmatrix}
		k_R c & \frac{k_\Omega c}{2} \\
		\frac{k_\Omega c}{2} & k_\Omega - c \lambda_M H
	\end{bmatrix} .
\end{gather*}
If \( c \) is chosen such that~\cref{eqn:c_bound} is satisfied then the matrices \( W_1, W_2 \) and \( M \) are positive definite.
This implies that $\mathcal{V}$ is positive definite and decrescent, and $\dot{\mathcal{V}}$ is negative semidefinite in the domain $D_3$. As such, the zero equilibrium is stable in the sense of Lyapunov, and all of the error variables are bounded. Furthermore, $\lim_{t\to\infty} \zeta=0$ according to the LaShalle-Yoshizawa theorem~\cite{khalil1996}.
